Analyzing how sociodemographic characteristics influence voting for the Liberal Party of Canada

### 2024-04-17 paper/paper.qmd
update

---
title: "Analysising how sociodemographic characteristics influence voting for the Liberal Party of Canada"
subtitle: "Decoding CES 2021 Vote Data by using A Multilevel Logistic Regression Model"
author: 
  - Yichen ji
thanks: "Code and data are available at: https://github.com/Selinayichenji/Canadian_election_prediction.git."
date: today
date-format: long
abstract: "This research utilized the 2021 Canadian Election Study and a multilevel logistic regression to examine how sociodemographic factors affect voting for the Liberal Party of Canada. It found that age, wealth, and education level significantly increase the likelihood of voting liberal, while children number, gender, and province do not. These results highlight the impact of life experiences and advantages on political ideologies and decision-making. Understanding these voting behaviors allows the Liberal Party of Canada to refine their electoral strategies, improving campaign effectiveness and voter engagement."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
link-citations: true
output:
  pdf_document: default

@@

The analysis reveals a

The remainder of this paper is structured as follows: the basic information contains data source, methodology, variables, measurements and the visualization of population structure and voting rate are in the @sec-data. The equation and interpretation of model are presented in @sec-models. The data and figure of coefficients are in the @sec-result. And in @sec-dis, we discuss what we have learned, our understanding of the world, the limitations, and the next steps of our research. @sec-A is for appendix and @sec-ref is listed all references in this paper.

@@

```
@tbl-liberal-3 shows the intercept for each province. The coefficients confirms our prediction from checking the percentage of voting for the liberal party in data section.

\newpage
# Discussion {#sec-dis}

@@

## What we learn about the world?
Through this study, we identified age, education level, economic status, and gender as significant factors influencing voter support for the Liberal Party. As age increases, voter support for the Liberal Party tends to increase as well, suggesting that older voters may prefer parties that promote stable policies and emphasize social welfare. Moreover, voters with higher education levels tend to support the Liberal Party, likely reflecting their approval of the party’s policies on education, environmental protection, and technological innovation. Similarly, the Liberal Party finds strong support among high-income groups, possibly due to the party's policies promoting economic development and business innovation.

Based on these findings, policymakers and the Liberal Party should consider developing specific policies to attract and maintain the support of these key voter groups. For instance, they could increase discussions on retirement insurance and health protection to appeal to older voters; continue supporting educational and research projects to solidify support among highly educated voters; and maintain support from high-income groups through tax incentives and business support strategies. Additionally, strengthening support for women's rights could also help to increase female voter turnout for the Liberal Party.

@@

# Additional data details

# Model details {#sec-model-details}

>>>>

---
title: "Analyzing how sociodemographic characteristics influence voting for the Liberal Party of Canada"
subtitle: "Decoding CES 2021 Vote Data by using A Multilevel Logistic Regression Model"
author: 
  - Yichen ji
thanks: "Code and data are available at: https://github.com/Selinayichenji/Canadian_election_prediction.git."
date: today
date-format: long
abstract: "This research utilized the 2021 Canadian Election Study and a multilevel logistic regression to examine how sociodemographic factors affect voting for the Liberal Party of Canada. It found that age, wealth, and education level significantly increase the likelihood of voting liberal, while children number, gender, and province do not. These results highlight the impact of life experiences and advantages on political ideologies and decision-making. Understanding these voting behaviors allows the Liberal Party of Canada to refine their electoral strategies, improving campaign effectiveness and voter engagement."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
link-citations: true
output:
  pdf_document: default
header-includes:
      - \usepackage{float}
      - \floatplacement{table}{H}

@@

The analysis reveals age, wealth, and education level have a positive relationship with voting liberal, while children number, gender, and province do not.

The remainder of this paper is structured as follows: the basic information contains data source, methodology, variables, measurements and the visualization of population structure and voting rate are in the @sec-data. The equation and interpretation of model are presented in @sec-models. The data and figure of coefficients are in the @sec-result. And in @sec-dis, we discuss what we have learned, our understanding of the world, the limitations, and the next steps of our research. @sec-A contains pp check and r-hat check for the model and @sec-ref is listed all references in this paper.

@@

```
@tbl-liberal-3 shows the intercept for each province. The coefficients align with the trend illustrated in @fig-vote-percent.

# Discussion {#sec-dis}

@@

## What we learn about the world?
Through this study, we have gained a nuanced understanding of the factors that influence voter support for the Liberal Party in Canada. As individuals age, their likelihood to support the Liberal Party increases, a trend particularly noticeable among older voters who have traditionally been Conservative supporters. This shift could be attributed to the Liberal Party's efforts in addressing issues pertinent to senior citizens, possibly influenced by recent events such as the COVID-19 vaccination rollout, where a significant percentage of senior Canadians have been vaccinated. The implication here is that policies that directly impact the well-being and concerns of older adults can be a powerful determinant of their voting preferences.

Furthermore, analysis of voting intentions suggests that gender plays a significant role among younger voters, where we see distinct patterns such as young men favoring the Conservative Party and young women showing stronger support for the NDP. Yet, for the Liberal Party, the support appears to be consistent across different segments, irrespective of age or gender. This consistency suggests that the Liberal Party's policies may have broad appeal across different demographics or that the party's stance on various issues is not polarizing enough to create a noticeable gender gap within its supporters.

Educational attainment is another influential factor, with those holding higher education levels likely resonating with the Liberal Party's emphasis on education, environmental protection, and technological innovation. These voters may view the Liberal Party as more aligned with their values and interests, particularly in policy areas that affect long-term societal progress.

Given these insights, it would be beneficial for policymakers and the Liberal Party to focus on developing and communicating policies that address the specific needs and concerns of these key voter groups. For the older demographic, continuing to provide and expand on healthcare, retirement insurance, and health protection will likely be vital. For the highly educated demographic, emphasizing the Liberal Party's commitment to advancing education and innovation could solidify support. Additionally, understanding the salient issues for different age and gender groups can help tailor communication and policy initiatives more effectively.

These observations are crucial for the Liberal Party if they are to maintain and strengthen their voter base, especially considering the evolving dynamics within the Canadian electorate.

@@
